---
title: 'Part 2: Adding Environmental Data'
author: "Veronica Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(auk)
library(lubridate)
library(sf)
library(gridExtra)
library(tidyverse)
library(rnaturalearth)
library(dplyr)

library(raster)
library(MODIS)
library(exactextractr)
library(viridis)
library(dggridR)
library(ranger)
library(scam)
library(PresenceAbsence)
library(verification)
library(ebirdst)
library(fields)
library(pscl)
library(MASS)

# Spatial packages
library(spdep)
library(rgdal)
library(tripack)
library(dbscan)
library(hglm)

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

## Part A: Loading GIS Data

```{r}
# Create a file for saving spatial data
gpkg_dir <- "data"
if (!dir.exists(gpkg_dir)) {
  dir.create(gpkg_dir)
}
f_ne <- file.path(gpkg_dir, "gis-data.gpkg")

# Load in manually downloaded BCR terrestrial shapefile for BCR 31
bcr <- file.path("D:", "bird folders", "White Ibis", "data", "BCR_Terrestrial_master_International.shp") %>% 
  read_sf() %>% 
  select(bcr_code = BCR, bcr_name = LABEL) %>% 
# Select BCR 31, Peninsular Florida
  filter(bcr_code %in% c(31))

# Load in manually downloaded Natural Earth GIS data

ne_land <- file.path("D:", "bird folders", "White Ibis", "data", "ne_50m_admin_0_countries_lakes.shp") %>%
  read_sf()

ne_land <- ne_land %>%
  st_set_precision(1e6) %>%
  st_union()

ne_country_lines <- file.path("D:", "bird folders", "White Ibis", "data", "ne_50m_admin_0_boundary_lines_land.shp") %>%
  read_sf()

ne_country_lines <- ne_country_lines %>%
  st_geometry()

ne_state_lines <- file.path("D:", "bird folders", "White Ibis", "data", "ne_50m_admin_1_states_provinces_lines.shp") %>%
  read_sf()

# Save for future use
unlink(f_ne)
write_sf(ne_land, f_ne, "ne_land")
write_sf(ne_country_lines, f_ne, "ne_country_lines")
write_sf(ne_state_lines, f_ne, "ne_state_lines")
write_sf(bcr, f_ne, "bcr")
```

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection

# Load in the white ibis data from Part 1
ibis <- read_csv("data/ebd_white_ibis.csv", show_col_types = FALSE)
```

## Part B: Preparing Environmental Covariates

Establish the spatial boundary for BCR 31

```{r}
# BCR 31 boundary
bcr <- read_sf("data/gis-data.gpkg", "bcr") %>% 
  filter(bcr_code == 31) %>% 
  # project to the native modis projection
  st_transform(crs = paste("+proj=sinu +lon_0=0 +x_0=0 +y_0=0",
                           "+a=6371007.181 +b=6371007.181 +units=m +no_defs"))
```

Get data for the MODIS tiles that cover BCR 31

```{r}
# Switch off spherical geometry
sf_use_s2(FALSE)

# Get tiles for BCR 31, the Floridian peninsula
tiles <- MODIS::getTile(bcr)
tiles@tile
```

Load this MODIS land cover data into R

```{r, warning = FALSE}
# Earliest year of ibis dataset
begin_year <- format(min(ibis$observation_date), "%Y.01.01")

# Last year of ibis dataset
end_year <- format(max(ibis$observation_date), "%Y.12.31")

# Download tiles and combine into a single raster for each year
tifs <- runGdal(product = "MCD12Q1", collection = "006", SDSstring = "01", 
                extent = bcr %>% st_buffer(dist = 10000), 
                begin = begin_year, end = end_year, 
                outDirPath = "data", job = "modis",
                MODISserverOrder = "LPDAAC") %>% 
  pluck("MCD12Q1.006") %>% 
  unlist()

# Rename the tifs to have better names
new_names <- format(as.Date(names(tifs)), "%Y") %>%
  sprintf("modis_mcd12q1_umd_%s.tif", .) %>%
  file.path(dirname(tifs), .)

file.rename(tifs, new_names)
```

```{r}
# Loading in the landcover data
landcover <- list.files("data/modis", "^modis_mcd12q1_umd",
                        full.names = TRUE) %>%
  stack()

# Label the layers with year
landcover <- names(landcover) %>%
  str_extract("(?<=modis_mcd12q1_umd_)[0-9]{4}") %>%
  paste0("y", .) %>%
  setNames(landcover, .)

# Extract the last year in the land cover data
max_lc_year <- names(landcover) %>%
  str_extract("[0-9]{4}") %>%
  as.integer() %>%
  max()
```

Establish neighborhood radius of 2.5 kilometers

```{r}
neighborhood_radius <- 5 * ceiling(max(res(landcover))) /2
```

Obtain the full set of distinct checklist locations by year, then create a neighborhood around each location by year

```{r}
bird_buff <- ibis %>%
  distinct(year = format(observation_date, "%Y"),
           locality_id, latitude, longitude) %>%
# In case the maximum ibis year is after the maximum land cover year available
# Extend the maximum land cover year data to all future years
  mutate(year_lc = if_else(as.integer(year) > max_lc_year,
                           as.character(max_lc_year), year),
         year_lc = paste0("y", year_lc)) %>%
# Convert to sf with crs = 4326
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
# Transform to the MODIS projection
  st_transform(crs = projection(landcover)) %>%
# Create an area around each point
  st_buffer(dist = neighborhood_radius) %>%
# Nest by year
  nest(data = c(year, locality_id, geometry))
```

Calculate the number of cells belonging to each land cover class for each checklist location, by year

```{r}
# Establish a function to summarize landcover data for all locations in a given year
calculate_pland <- function(yr, regions, lc) {
  locs <- st_set_geometry(regions, NULL)
  exact_extract(lc[[yr]], regions, progress = FALSE) %>%
    map(~ count(., landcover = value)) %>%
    tibble(locs, data = .) %>%
    unnest(data)
}

# Iterate over all years, extracting the land cover values for all locations in each year
lc_extract <- bird_buff %>%
  mutate(pland = map2(year_lc, data, calculate_pland, lc = landcover)) %>%
  select(pland) %>%
  unnest(cols = pland)
```

Calculate the proportion of land cover for each location and land cover class, by year

```{r}
pland <- lc_extract %>%
# Calculate proportions
  group_by(locality_id, year) %>%
  mutate(pland = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  filter(!is.na(landcover))
```

Change covariate names and turn the data into a wide format

```{r}
lc_names <- tibble(landcover = 0:15,
                   lc_name = c("pland_00_water", 
                               "pland_01_evergreen_needleleaf", 
                               "pland_02_evergreen_broadleaf", 
                               "pland_03_deciduous_needleleaf", 
                               "pland_04_deciduous_broadleaf", 
                               "pland_05_mixed_forest",
                               "pland_06_closed_shrubland", 
                               "pland_07_open_shrubland", 
                               "pland_08_woody_savanna", 
                               "pland_09_savanna", 
                               "pland_10_grassland", 
                               "pland_11_wetland", 
                               "pland_12_cropland", 
                               "pland_13_urban", 
                               "pland_14_mosaic", 
                               "pland_15_barren"))

pland <- pland %>%
  inner_join(lc_names, by = "landcover") %>%
  arrange(landcover) %>%
  select(-landcover)

# As I transform to the wide format, I'll fill in any implicit missing values with 0. 
pland <- pland %>%
  pivot_wider(names_from = lc_name,
              values_from = pland,
              values_fill = list(pland = 0))
```

Manually re-create pland 03, since it is "lost" / missing for an unknown reason

```{r}
pland <- pland %>%
  mutate(pland_03_deciduous_needleleaf = 1 - (pland_00_water
                               + pland_01_evergreen_needleleaf 
                               + pland_02_evergreen_broadleaf
                               + pland_04_deciduous_broadleaf 
                               + pland_05_mixed_forest
                               + pland_06_closed_shrubland
                               + pland_07_open_shrubland 
                               + pland_08_woody_savanna 
                               + pland_09_savanna
                               + pland_10_grassland
                               + pland_11_wetland 
                               + pland_12_cropland 
                               + pland_13_urban
                               + pland_14_mosaic 
                               + pland_15_barren))
```

Save the land cover file

```{r}
write_csv(pland, "data/modis_pland_location_year.csv")
```

Load manually downloaded elevation data, then crop to our study area

```{r}
# Load in the data
elev <- raster("data/elevation_1KMmd_GMTEDmd.tif")

# Crop to BCR 31, including a buffer for "wiggle room"
elev <- bcr %>%
  st_buffer(dist = 10000) %>%
  st_transform(crs = projection(elev)) %>%
  crop(elev, .) %>%
  projectRaster(crs = projection(landcover))
```

Extract the elevation values that correspond to each area, then use these values to calculate mean and standard deviation of elevation in each area

```{r}
bird_buff_no_year <- ibis %>%
  distinct(locality_id, latitude, longitude) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = projection(elev)) %>%
  st_buffer(dist = neighborhood_radius)

locs <- st_set_geometry(bird_buff_no_year, NULL) %>%
  mutate(id = row_number())

elev_checklists <- exact_extract(elev, bird_buff_no_year, progress = FALSE) %>%
  map_dfr(~ tibble(elevation_mean = mean(.$value, na.rm = TRUE),
                   elevation_median = median(.$value, na.rm = TRUE),
                   elevation_sd = sd(.$value, na.rm = TRUE),
                   elevation_iqr = IQR(.$value, na.rm = TRUE))) %>%

  bind_cols(locs, .)
```

Merge the environmental covariates together and save the dataset

```{r}
# Merge the land cover and elevation covariates
pland_elev_checklist <- inner_join(pland, elev_checklists, by = "locality_id")

# Save the data
write_csv(pland_elev_checklist, "data/pland_elev_location_year.csv")
```

## Part C: Merging the Data

## Prepare the data

```{r}
# Establish the output directory

if (!dir.exists("output")) {
  dir.create("output")
}
```

```{r}
# Read in all ten sets of eBird data and the environmental covariates
# Establish the variable `year` for all datasets for the purposes of merging the data

white.ibis <- read_csv("data/ebd_white_ibis.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))
```

```{r}
glossy.ibis <- read_csv("data/ebd_glossy_ibis.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

roseate <- read_csv("data/ebd_spoonbill.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

great.egret <- read_csv("data/ebd_great_egret.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

snowy.egret <- read_csv("data/ebd_snowy_egret.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

cattle.egret <- read_csv("data/ebd_cattle_egret.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

# Note: "gbh" is Great Blue Heron

gbh <- read_csv("data/ebd_gbh.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

# Note: "lbh" is Little Blue Heron

lbh <- read_csv("data/ebd_lbh.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

tri.heron <- read_csv("data/ebd_tri_heron.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))

green.heron <- read_csv("data/ebd_green_heron.csv", show_col_types = FALSE) %>%
  mutate(year = year(observation_date))
```

```{r}
habitat <- read_csv("data/pland_elev_location_year.csv", show_col_types = FALSE) %>%
  mutate(year = as.integer(year))
```

```{r}
# Troubleshooting: this should print TRUE
nrow(habitat) == 21751
```

## White ibis

```{r}
# Join the datasets 
white.ibis.habitat <- inner_join(white.ibis, habitat, by = c("locality_id", "year"))
```

```{r}
# Troubleshooting
nrow(white.ibis.habitat) == 51496

nrow(filter(white.ibis.habitat, year == 2016)) == 3859

nrow(filter(white.ibis.habitat, year == 2017)) == 5181
```


```{r}
# Remove NA counts
white.ibis.abundance <- white.ibis.habitat %>% filter(!is.na(observation_count))
```

```{r}
# Rename two variables
white.ibis.abundance$distance_traveled_km <- white.ibis.abundance$effort_distance_km

white.ibis.abundance$bird_count <- white.ibis.abundance$observation_count

# Filter training data and save the file
white.ibis.2016 <- white.ibis.abundance %>% filter(year == 2016)

write_csv(white.ibis.2016, "data/ebd_white_ibis_2016.csv", na = "")

# Filter test data and save the file
white.ibis.2017 <- white.ibis.abundance %>% filter(year == 2017)

write_csv(white.ibis.2017, "data/ebd_white_ibis_2017.csv", na = "")
```

## Glossy ibis

```{r}
# Join the datasets 
glossy.ibis.habitat <- inner_join(glossy.ibis, habitat, by = c("locality_id", "year"))

# Remove NA counts
glossy.ibis.abundance <- glossy.ibis.habitat %>% filter(!is.na(observation_count))

# Rename two variables
glossy.ibis.abundance$distance_traveled_km <- glossy.ibis.abundance$effort_distance_km

glossy.ibis.abundance$bird_count <- glossy.ibis.abundance$observation_count

# Filter training data and save the file
glossy.ibis.2016 <- glossy.ibis.abundance %>% filter(year == 2016)

write_csv(glossy.ibis.2016, "data/ebd_glossy_ibis_2016.csv", na = "")

# Filter test data and save the file
glossy.ibis.2017 <- glossy.ibis.abundance %>% filter(year == 2017)

write_csv(glossy.ibis.2017, "data/ebd_glossy_ibis_2017.csv", na = "")
```

## Roseate spoonbill

```{r}
# Join the datasets 
roseate.habitat <- inner_join(roseate, habitat, by = c("locality_id", "year"))

# Remove NA counts
roseate.abundance <- roseate.habitat %>% filter(!is.na(observation_count))

# Rename two variables
roseate.abundance$distance_traveled_km <- roseate.abundance$effort_distance_km

roseate.abundance$bird_count <- roseate.abundance$observation_count

# Filter training data and save the file
roseate.2016 <- roseate.abundance %>% filter(year == 2016)

write_csv(roseate.2016, "data/ebd_spoonbill_2016.csv", na = "")

# Filter test data and save the file
roseate.2017 <- roseate.abundance %>% filter(year == 2017)

write_csv(roseate.2017, "data/ebd_spoonbill_2017.csv", na = "")
```

## Great egret

```{r}
# Join the datasets 
great.egret.habitat <- inner_join(great.egret, habitat, by = c("locality_id", "year"))

# Remove NA counts
great.egret.abundance <- great.egret.habitat %>% filter(!is.na(observation_count))

# Rename two variables
great.egret.abundance$distance_traveled_km <- great.egret.abundance$effort_distance_km

great.egret.abundance$bird_count <- great.egret.abundance$observation_count

# Filter training data and save the file
great.egret.2016 <- great.egret.abundance %>% filter(year == 2016)

write_csv(great.egret.2016, "data/ebd_great_egret_2016.csv", na = "")

# Filter test data and save the file
great.egret.2017 <- great.egret.abundance %>% filter(year == 2017)

write_csv(great.egret.2017, "data/ebd_great_egret_2017.csv", na = "")
```

## Cattle egret

```{r}
# Join the datasets 
cattle.egret.habitat <- inner_join(cattle.egret, habitat, by = c("locality_id", "year"))

# Remove NA counts
cattle.egret.abundance <- cattle.egret.habitat %>% filter(!is.na(observation_count))

# Rename two variables
cattle.egret.abundance$distance_traveled_km <- cattle.egret.abundance$effort_distance_km

cattle.egret.abundance$bird_count <- cattle.egret.abundance$observation_count

# Filter training data and save the file
cattle.egret.2016 <- cattle.egret.abundance %>% filter(year == 2016)

write_csv(cattle.egret.2016, "data/ebd_cattle_egret_2016.csv", na = "")

# Filter test data and save the file
cattle.egret.2017 <- cattle.egret.abundance %>% filter(year == 2017)

write_csv(cattle.egret.2017, "data/ebd_cattle_egret_2017.csv", na = "")
```

## Snowy egret

```{r}
# Join the datasets 
snowy.egret.habitat <- inner_join(snowy.egret, habitat, by = c("locality_id", "year"))

# Remove NA counts
snowy.egret.abundance <- snowy.egret.habitat %>% filter(!is.na(observation_count))

# Rename two variables
snowy.egret.abundance$distance_traveled_km <- snowy.egret.abundance$effort_distance_km

snowy.egret.abundance$bird_count <- snowy.egret.abundance$observation_count

# Filter training data and save the file
snowy.egret.2016 <- snowy.egret.abundance %>% filter(year == 2016)

write_csv(snowy.egret.2016, "data/ebd_snowy_egret_2016.csv", na = "")

# Filter test data and save the file
snowy.egret.2017 <- snowy.egret.abundance %>% filter(year == 2017)

write_csv(snowy.egret.2017, "data/ebd_snowy_egret_2017.csv", na = "")
```

## Great blue heron

```{r}
# Join the datasets 
gbh.habitat <- inner_join(gbh, habitat, by = c("locality_id", "year"))

# Remove NA counts
gbh.abundance <- gbh.habitat %>% filter(!is.na(observation_count))

# Rename two variables
gbh.abundance$distance_traveled_km <- gbh.abundance$effort_distance_km

gbh.abundance$bird_count <- gbh.abundance$observation_count

# Filter training data and save the file
gbh.2016 <- gbh.abundance %>% filter(year == 2016)

write_csv(gbh.2016, "data/ebd_gbh_2016.csv", na = "")

# Filter test data and save the file
gbh.2017 <- gbh.abundance %>% filter(year == 2017)

write_csv(gbh.2017, "data/ebd_gbh_2017.csv", na = "")
```

## Little blue heron

```{r}
# Join the datasets 
lbh.habitat <- inner_join(lbh, habitat, by = c("locality_id", "year"))

# Remove NA counts
lbh.abundance <- lbh.habitat %>% filter(!is.na(observation_count))

# Rename two variables
lbh.abundance$distance_traveled_km <- lbh.abundance$effort_distance_km

lbh.abundance$bird_count <- lbh.abundance$observation_count

# Filter training data and save the file
lbh.2016 <- lbh.abundance %>% filter(year == 2016)

write_csv(lbh.2016, "data/ebd_lbh_2016.csv", na = "")

# Filter test data and save the file
lbh.2017 <- lbh.abundance %>% filter(year == 2017)

write_csv(lbh.2017, "data/ebd_lbh_2017.csv", na = "")
```

## Tricolored heron

```{r}
# Join the datasets 
tri.heron.habitat <- inner_join(tri.heron, habitat, by = c("locality_id", "year"))

# Remove NA counts
tri.heron.abundance <- tri.heron.habitat %>% filter(!is.na(observation_count))

# Rename two variables
tri.heron.abundance$distance_traveled_km <- tri.heron.abundance$effort_distance_km

tri.heron.abundance$bird_count <- tri.heron.abundance$observation_count

# Filter training data and save the file
tri.heron.2016 <- tri.heron.abundance %>% filter(year == 2016)

write_csv(tri.heron.2016, "data/ebd_tri_heron_2016.csv", na = "")

# Filter test data and save the file
tri.heron.2017 <- tri.heron.abundance %>% filter(year == 2017)

write_csv(tri.heron.2017, "data/ebd_tri_heron_2017.csv", na = "")
```

## Green heron

```{r}
# Join the datasets 
green.heron.habitat <- inner_join(green.heron, habitat, by = c("locality_id", "year"))

# Remove NA counts
green.heron.abundance <- green.heron.habitat %>% filter(!is.na(observation_count))

# Rename two variables
green.heron.abundance$distance_traveled_km <- green.heron.abundance$effort_distance_km

green.heron.abundance$bird_count <- green.heron.abundance$observation_count

# Filter training data and save the file
green.heron.2016 <- green.heron.abundance %>% filter(year == 2016)

write_csv(green.heron.2016, "data/ebd_green_heron_2016.csv", na = "")

# Filter test data and save the file
green.heron.2017 <- green.heron.abundance %>% filter(year == 2017)

write_csv(green.heron.2017, "data/ebd_green_heron_2017.csv", na = "")
```
