---
title: 'Part 7: Producing Effect Displays'
author: "Veronica Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)
library(MODIS)
library(exactextractr)
library(viridis)
library(tidyverse)

library(dggridR)
library(lubridate)
library(ranger)
library(scam)
library(PresenceAbsence)
library(verification)
library(ebirdst)
library(fields)
library(gridExtra)
library(pscl)
library(MASS)
library(ggpubr)

# Spatial packages
library(spdep)
library(rgdal)
library(tripack)
library(dbscan)
library(hglm)

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

# Part A: Effect Displays

## Load the data and models

```{r}
# Load training data
ibis <- read_csv("data/ebd_white_ibis_train.csv", show_col_types = FALSE)

# Load test data
ibis.2017 <- read_csv("data/ebd_white_ibis_2017.csv", show_col_types = FALSE)

# Load GAM
load("white_ibis_gam.model")

# Load HGAM
load("white_ibis_hgam.model")
```

## Histograms for selected covariates

### Time Observations Started

```{r}
ggplot(ibis, aes(x = time_observations_started)) +
  geom_histogram()
```

We will set the range of "time observations started" values in its effect display from the 0.01 percentile to the 0.99 percentile, as there are both extreme low and extreme high values for this covariate.

### Wetland

```{r}
ggplot(ibis, aes(x = pland_11_wetland)) +
  geom_histogram()
```

There are extreme values of wetland on the right-hand side of the distribution. We will set the range of the effect display for wetland from its minimum value to its 0.99 percentile. 

### Urban

```{r}
ggplot(ibis, aes(x = pland_13_urban)) +
  geom_histogram()
```

The distribution of the urban covariate does not show any unusual extreme low or high values. The range of the effect display for urban will run from the covariate's minimum to its maximum.

## Effect displays for GAM

### Time Observations Started / Hours Since Midnight

```{r}
tos.01 <- as.numeric(quantile(ibis$time_observations_started, probs = 0.01))
tos.99 <- as.numeric(quantile(ibis$time_observations_started, probs = 0.99))
tos_seq <- seq(tos.01, tos.99, length.out = 300)

tos_data <- ibis %>%
  select(duration_minutes, distance_traveled_km, number_observers, pland_02_evergreen_broadleaf, pland_04_deciduous_broadleaf, pland_08_woody_savanna, pland_10_grassland, pland_11_wetland, pland_12_cropland, pland_13_urban, pland_14_mosaic, elevation_mean, elevation_sd) %>%
  summarize_all(median, na.ram = TRUE) %>%
  ungroup() %>%
  cbind(time_observations_started = tos_seq)
```

```{r}
# Then predict counts using GAM model
tos_link <- predict(qpois.gam, newdata = tos_data,
                    type = "link",
                    se.fit = TRUE)

tos_predictions <- as_tibble(tos_link) %>% transmute(time_observations_started = tos_seq,
            pred_counts = qpois.gam$family$linkinv(tos_link$fit))
```

```{r}
# Generate the effect display plot
plot1 <- tos_predictions %>%
ggplot(aes(x = time_observations_started, y = pred_counts)) +
  geom_line() +
  labs(x = "Hours since midnight, GAM") +
  theme(axis.title.y = element_blank()) +
  ylim(0, 2.25)
```

### Wetland

```{r}
# Create a vector of covariates with a range of "pland 11 (wetland)" values
# Range is from the minimum of pland 11 to the 0.99 percentile of pland 11
min.11 <- min(ibis$pland_11_wetland)
percentile.11 <- as.numeric(quantile(ibis$pland_11_wetland, probs = 0.99))

pland_11_seq <- seq(min.11, percentile.11, length.out = 300)

# Create a dataframe: all other covariates will have median values
pland_11_data <- ibis %>%
  select(time_observations_started, duration_minutes, distance_traveled_km, number_observers, pland_02_evergreen_broadleaf, pland_04_deciduous_broadleaf, pland_08_woody_savanna, pland_10_grassland, pland_12_cropland, pland_13_urban, pland_14_mosaic, elevation_mean, elevation_sd) %>%
  summarize_all(median, na.ram = TRUE) %>%
  ungroup() %>%
  cbind(pland_11_wetland = pland_11_seq)
```

```{r}
# Predict counts using GAM
pland_11_link <- predict(qpois.gam, newdata = pland_11_data,
                    type = "link",
                    se.fit = TRUE)

pland_11_predictions <- as_tibble(pland_11_link) %>% transmute(pland_11_wetland = pland_11_seq,
            pred_counts = qpois.gam$family$linkinv(pland_11_link$fit))
```

```{r}
# Generate the effect display plot
plot7 <- pland_11_predictions %>%
ggplot(aes(x = pland_11_wetland, y = pred_counts)) +
  geom_line() +
  labs(x = "Proportion of wetland terrain, GAM") +
  theme(axis.title.y = element_blank()) +
  ylim(0, 4)
```

### Urban

```{r}
# Create a vector of covariates with a range of "pland 13 (urban)" values
# Range is from the minimum of pland 13 to the maximum value of pland 13
min.13 <- min(ibis$pland_13_urban)
max.13 <- max(ibis$pland_13_urban)

pland_13_seq <- seq(min.13, max.13, length.out = 300)

# Create a dataframe: all other covariates will have median values
pland_13_data <- ibis %>%
  select(time_observations_started, duration_minutes, distance_traveled_km, number_observers, pland_02_evergreen_broadleaf, pland_04_deciduous_broadleaf, pland_08_woody_savanna, pland_10_grassland, pland_11_wetland, pland_12_cropland, pland_14_mosaic, elevation_mean, elevation_sd) %>%
  summarize_all(median, na.ram = TRUE) %>%
  ungroup() %>%
  cbind(pland_13_urban = pland_13_seq)
```

```{r}
# Predict counts using GAM
pland_13_link <- predict(qpois.gam, newdata = pland_13_data,
                    type = "link",
                    se.fit = TRUE)

pland_13_predictions <- as_tibble(pland_13_link) %>% transmute(pland_13_urban = pland_13_seq,
            pred_counts = qpois.gam$family$linkinv(pland_13_link$fit))
```

```{r}
# Generate the effect display plot
plot8 <- pland_13_predictions %>%
ggplot(aes(x = pland_13_urban, y = pred_counts)) +
  geom_line() +
  labs(x = "Proportion of urban terrain, GAM") +
  theme(axis.title.y = element_blank()) +
  ylim(0, 2.25)
```

## Effect displays for HGAM

### Time Observations Started / Hours Since Midnight

```{r}
# Save a model matrix for tos_data
tos.matrix <- predict(qpois.gam, newdata = tos_data, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- hgam.soi$fixef
hgam.coef <- as.matrix(hgam.coef)

# Generate "link" predictions
tos_link <- (tos.matrix %*% hgam.coef)

# Prepare the tibble
time_observations_started <- tos_seq
tos_pred <- exp(tos_link)

tos_predictions <- tibble(time_observations_started, tos_pred)
```

```{r}
# Generate the effect display plot
plot01 <- tos_predictions %>%
ggplot(aes(x = time_observations_started, y = tos_pred)) +
  geom_line() +
  labs(x = "Hours since midnight, HGAM") +
  theme(axis.title.y = element_blank()) +
  ylim(0, 2.25)
```

### Wetland

```{r}
# Save a model matrix for pland_11_data
pland11.matrix <- predict(qpois.gam, newdata = pland_11_data, type = "lpmatrix")

# Generate "link" predictions
pland_11_link <- (pland11.matrix %*% hgam.coef)

# Prepare the tibble
pland_11_wetland <- pland_11_seq

pland_11_pred <- exp(pland_11_link)

pland_11_predictions <- tibble(pland_11_wetland, pland_11_pred)
```

```{r}
# Generate the effect display plot
plot07 <- pland_11_predictions %>%
ggplot(aes(x = pland_11_wetland, y = pland_11_pred)) +
  geom_line() +
  labs(x = "Proportion of wetland terrain, HGAM") +
  theme(axis.title.y = element_blank()) +
  ylim(0, 4)
```

### Urban

```{r}
# Save a model matrix for pland_13_data
pland13.matrix <- predict(qpois.gam, newdata = pland_13_data, type = "lpmatrix")

# Generate "link" predictions
pland_13_link <- (pland13.matrix %*% hgam.coef)

# Prepare the tibble
pland_13_urban <- pland_13_seq

pland_13_pred <- exp(pland_13_link)

pland_13_predictions <- tibble(pland_13_urban, pland_13_pred)
```

```{r}
# Generate the effect display plot
plot08 <- pland_13_predictions %>%
ggplot(aes(x = pland_13_urban, y = pland_13_pred)) +
  geom_line() +
  labs(x = "Proportion of urban terrain, HGAM") +
  theme(axis.title.y = element_blank()) +
  ylim(0, 2.25)
```

## Generate final image

```{r}
grob <- arrangeGrob(plot1, plot7, plot8, plot01, plot07, plot08, ncol = 3, nrow = 2, top = text_grob("Effect of Selected Covariates on White Ibis Relative Abundance", size = 16), left = textGrob("Relative abundance", rot = 90))

ggsave(file="effect_displays_final.png", grob, width = 10, height = 4)
```
