---
title: 'Part 8: Producing Maps'
author: "Veronica Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(auk)
library(lubridate)
library(sf)
library(gridExtra)
library(tidyverse)
library(rnaturalearth)
library(dplyr)

library(raster)
library(MODIS)
library(exactextractr)
library(viridis)
library(dggridR)
library(ranger)
library(scam)
library(PresenceAbsence)
library(verification)
library(ebirdst)
library(fields)
library(pscl)
library(MASS)

# Spatial packages
library(spdep)
library(rgdal)
library(tripack)
library(dbscan)
library(hglm)

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection

# Load in the white ibis data from Part 1
ibis <- read_csv("data/ebd_white_ibis.csv", show_col_types = FALSE)
```

## From Part 2.B: Preparing Environmental Covariates

Establish the spatial boundary for BCR 31

```{r}
# BCR 31 boundary
bcr <- read_sf("data/gis-data.gpkg", "bcr") %>% 
  filter(bcr_code == 31) %>% 
  # project to the native modis projection
  st_transform(crs = paste("+proj=sinu +lon_0=0 +x_0=0 +y_0=0",
                           "+a=6371007.181 +b=6371007.181 +units=m +no_defs"))
```

Get data for the MODIS tiles that cover BCR 31

```{r}
# Switch off spherical geometry
sf_use_s2(FALSE)

# Get tiles for BCR 31, the Floridian peninsula
tiles <- MODIS::getTile(bcr)
tiles@tile
```

Load this MODIS land cover data into R

```{r, warning = FALSE}
# Earliest year of ibis dataset
begin_year <- format(min(ibis$observation_date), "%Y.01.01")

# Last year of ibis dataset
end_year <- format(max(ibis$observation_date), "%Y.12.31")

# Download tiles and combine into a single raster for each year
tifs <- runGdal(product = "MCD12Q1", collection = "006", SDSstring = "01", 
                extent = bcr %>% st_buffer(dist = 10000), 
                begin = begin_year, end = end_year, 
                outDirPath = "data", job = "modis",
                MODISserverOrder = "LPDAAC") %>% 
  pluck("MCD12Q1.006") %>% 
  unlist()

# Rename the tifs to have better names
new_names <- format(as.Date(names(tifs)), "%Y") %>%
  sprintf("modis_mcd12q1_umd_%s.tif", .) %>%
  file.path(dirname(tifs), .)

file.rename(tifs, new_names)
```

```{r}
# Loading in the landcover data
landcover <- list.files("data/modis", "^modis_mcd12q1_umd",
                        full.names = TRUE) %>%
  stack()

# Label the layers with year
landcover <- names(landcover) %>%
  str_extract("(?<=modis_mcd12q1_umd_)[0-9]{4}") %>%
  paste0("y", .) %>%
  setNames(landcover, .)

# Extract the last year in the land cover data
max_lc_year <- names(landcover) %>%
  str_extract("[0-9]{4}") %>%
  as.integer() %>%
  max()
```

Establish neighborhood radius of 2.5 kilometers

```{r}
neighborhood_radius <- 5 * ceiling(max(res(landcover))) /2
```

Obtain the full set of distinct checklist locations by year, then create a neighborhood around each location by year

```{r}
bird_buff <- ibis %>%
  distinct(year = format(observation_date, "%Y"),
           locality_id, latitude, longitude) %>%
# In case the maximum ibis year is after the maximum land cover year available
# Extend the maximum land cover year data to all future years
  mutate(year_lc = if_else(as.integer(year) > max_lc_year,
                           as.character(max_lc_year), year),
         year_lc = paste0("y", year_lc)) %>%
# Convert to sf with crs = 4326
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
# Transform to the MODIS projection
  st_transform(crs = projection(landcover)) %>%
# Create an area around each point
  st_buffer(dist = neighborhood_radius) %>%
# Nest by year
  nest(data = c(year, locality_id, geometry))
```

Calculate the number of cells belonging to each land cover class for each checklist location, by year

```{r}
# Establish a function to summarize landcover data for all locations in a given year
calculate_pland <- function(yr, regions, lc) {
  locs <- st_set_geometry(regions, NULL)
  exact_extract(lc[[yr]], regions, progress = FALSE) %>%
    map(~ count(., landcover = value)) %>%
    tibble(locs, data = .) %>%
    unnest(data)
}

# Iterate over all years, extracting the land cover values for all locations in each year
lc_extract <- bird_buff %>%
  mutate(pland = map2(year_lc, data, calculate_pland, lc = landcover)) %>%
  select(pland) %>%
  unnest(cols = pland)
```

Calculate the proportion of land cover for each location and land cover class, by year

```{r}
pland <- lc_extract %>%
# Calculate proportions
  group_by(locality_id, year) %>%
  mutate(pland = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  filter(!is.na(landcover))
```

Change covariate names and turn the data into a wide format

```{r}
lc_names <- tibble(landcover = 0:15,
                   lc_name = c("pland_00_water", 
                               "pland_01_evergreen_needleleaf", 
                               "pland_02_evergreen_broadleaf", 
                               "pland_03_deciduous_needleleaf", 
                               "pland_04_deciduous_broadleaf", 
                               "pland_05_mixed_forest",
                               "pland_06_closed_shrubland", 
                               "pland_07_open_shrubland", 
                               "pland_08_woody_savanna", 
                               "pland_09_savanna", 
                               "pland_10_grassland", 
                               "pland_11_wetland", 
                               "pland_12_cropland", 
                               "pland_13_urban", 
                               "pland_14_mosaic", 
                               "pland_15_barren"))

pland <- pland %>%
  inner_join(lc_names, by = "landcover") %>%
  arrange(landcover) %>%
  select(-landcover)

# As I transform to the wide format, I'll fill in any implicit missing values with 0. 
pland <- pland %>%
  pivot_wider(names_from = lc_name,
              values_from = pland,
              values_fill = list(pland = 0))
```

Manually re-create pland 03, since it is "lost" / missing for an unknown reason

```{r}
pland <- pland %>%
  mutate(pland_03_deciduous_needleleaf = 1 - (pland_00_water
                               + pland_01_evergreen_needleleaf 
                               + pland_02_evergreen_broadleaf
                               + pland_04_deciduous_broadleaf 
                               + pland_05_mixed_forest
                               + pland_06_closed_shrubland
                               + pland_07_open_shrubland 
                               + pland_08_woody_savanna 
                               + pland_09_savanna
                               + pland_10_grassland
                               + pland_11_wetland 
                               + pland_12_cropland 
                               + pland_13_urban
                               + pland_14_mosaic 
                               + pland_15_barren))
```

Load manually downloaded elevation data, then crop to our study area

```{r}
# Load in the data
elev <- raster("data/elevation_1KMmd_GMTEDmd.tif")

# Crop to BCR 31, including a buffer for "wiggle room"
elev <- bcr %>%
  st_buffer(dist = 10000) %>%
  st_transform(crs = projection(elev)) %>%
  crop(elev, .) %>%
  projectRaster(crs = projection(landcover))
```

Extract the elevation values that correspond to each area, then use these values to calculate mean and standard deviation of elevation in each area

```{r}
bird_buff_no_year <- ibis %>%
  distinct(locality_id, latitude, longitude) %>%
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = projection(elev)) %>%
  st_buffer(dist = neighborhood_radius)

locs <- st_set_geometry(bird_buff_no_year, NULL) %>%
  mutate(id = row_number())

elev_checklists <- exact_extract(elev, bird_buff_no_year, progress = FALSE) %>%
  map_dfr(~ tibble(elevation_mean = mean(.$value, na.rm = TRUE),
                   elevation_median = median(.$value, na.rm = TRUE),
                   elevation_sd = sd(.$value, na.rm = TRUE),
                   elevation_iqr = IQR(.$value, na.rm = TRUE))) %>%

  bind_cols(locs, .)
```

Merge the environmental covariates together and save the dataset

```{r}
# Merge the land cover and elevation covariates
pland_elev_checklist <- inner_join(pland, elev_checklists, by = "locality_id")
```

## Prediction Surface

We will create a "prediction surface" for BCR 31 using the MODIS land cover data from 2016.

First, we will create a template raster with cells of the same size as the neighborhoods we previously defined. We will also assign the value 1 to all cells within BCR 31, leaving the cells outside the Floridian peninsula empty, and then remove the empty cells.

```{r}
agg_factor <- round(2 * neighborhood_radius / res(landcover))

r <- raster(landcover) %>%
  aggregate(agg_factor)

r <- bcr %>%
  st_transform(crs = projection(r)) %>%
  rasterize(r, field = 1) %>%
  trim()

r <- writeRaster(r, filename = "data/prediction_surface.tif", overwrite = TRUE)
```

For each cell, we will calculate the PLAND metrics using the same approach we used previously for the observation neighborhoods.

First, we will get cell centers and create neighborhoods:

```{r}
r_centers <- rasterToPoints(r, spatial = TRUE) %>%
  st_as_sf() %>%
  transmute(id = row_number())

r_cells <- st_buffer(r_centers, dist = neighborhood_radius)
```

Next, we will extract landcover values within each neighborhood, using the year of 2016.

```{r}
lc_extract_pred <- landcover[[paste0("y", 2016)]] %>%
  exact_extract(r_cells, progress = FALSE) %>%
  map(~ count(., landcover = value)) %>%
  tibble(id = r_cells$id, data = .) %>%
  unnest(data)
```

Following the above, we will calculate PLAND values across the prediction surface.

```{r}
pland_pred <- lc_extract_pred %>%
  count(id, landcover) %>%
  group_by(id) %>%
  mutate(pland = n / sum(n)) %>%
  ungroup() %>%
  select(-n) %>%
  filter(!is.na(landcover))
```

As before, we will change the names to be more descriptive.

```{r}
pland_pred <- pland_pred %>%
  inner_join(lc_names, by = "landcover") %>%
  arrange(landcover) %>%
  select(-landcover)
```

As before, we will transform to a wide data format and fill in implicit missing values with 0s.

```{r}
pland_pred <- pland_pred %>%
  pivot_wider(names_from = lc_name,
              values_from = pland,
              values_fill = list(pland = 0)) %>%
  mutate(year = 2016) %>%
  select(id, year, everything())
```

Now we will join in the coordinates.

```{r}
pland_coords <- st_transform(r_centers, crs = 4326) %>%
  st_coordinates() %>%
  as.data.frame() %>%
  cbind(id = r_centers$id, .) %>%
  rename(longitude = X, latitude = Y) %>%
  inner_join(pland_pred, by = "id")
```

We will now prepare elevation covariates for the prediction surface.

```{r}
elev_pred <- exact_extract(elev, r_cells, progress = FALSE) %>%
  map_dfr(~ tibble(elevation_mean = mean(.$value, na.rm = TRUE),
                   elevation_median = median(.$value, na.rm = TRUE),
                   elevation_sd = sd(.$value, na.rm = TRUE),
                   elevation_iqr = IQR(.$value, na.rm = TRUE))) %>%
# Join to lookup table to get locality_id: Not sure what that means
  bind_cols(st_drop_geometry(r_cells), .)
```

We will combine and save the PLAND and elevation covariates for the prediction surface.

```{r}
# Prediction surface covariates
pland_elev_pred <- inner_join(pland_coords, elev_pred, by = "id")

write_csv(pland_elev_pred, "data/pland_elev_prediction_surface.csv")
```

Finally, we will add a standard suite of effort covariates to every cell. In particular: the day is 6/15/16, time started is 12:00 p.m., duration is 60 minutes, distance is one kilometer, number of observers is one, and protocol type is Traveling.

```{r}
# Add a standard suite of effort covariates to the prediction surface
pred_surface_eff <- pland_elev_pred  %>% 
  mutate(day_of_year = yday(ymd(str_glue("2016-06-15"))),
         time_observations_started = 12,
         duration_minutes = 60,
         distance_traveled_km = 1,
         number_observers = 1,
         protocol_type = "Traveling")
```

## Load necessary GIS data

```{r}
map_proj <- st_crs('ESRI:102003')

ne_land <- read_sf("data/gis-data.gpkg", "ne_land") %>%
  st_set_crs(4326) %>%
  st_transform(crs = map_proj) %>%
  st_geometry()

bcr <- read_sf("data/gis-data.gpkg", "bcr") %>%
  st_set_crs(4326) %>%
  st_transform(crs = map_proj) %>%
  st_geometry()

ne_country_lines <- read_sf("data/gis-data.gpkg", "ne_country_lines") %>%
  st_set_crs(4326) %>%
  st_transform(crs = map_proj) %>%
  st_geometry()

ne_state_lines <- read_sf("data/gis-data.gpkg", "ne_state_lines") %>%
  st_set_crs(4326) %>%
  st_transform(crs = map_proj) %>%
  st_geometry()
```

## Wetland Plot

We will now make a plot of the proportion of wetland terrain across BCR 31.

```{r}
# Use if saving the plot to PDF
# pdf("wetland.pdf")

wetland_cover <- pland_coords %>%
# Convert to spatial features
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
  st_transform(crs = projection(r)) %>%
# Rasterize points
  rasterize(r, field = "pland_11_wetland") %>%
# Project for mapping
  projectRaster(crs = st_crs('ESRI:102003')$proj4string, method = "ngb") %>%
# Trim the empty edges of the raster
  trim()

# png(file = "pland_wetland.png")

par(mar = c(0.25, 0.25, 0.25, 0.25))

plot(bcr, col = NA, border = NA)
plot(ne_land, col = "#dddddd", border = "#888888", lwd = 0.75, add = TRUE)
    
plot(wetland_cover, axes = FALSE, box = FALSE, col = mako(20, direction = -1), add = TRUE)

  # borders
plot(bcr, border = "#000000", col = NA, lwd = 1.5, add = TRUE)
plot(ne_state_lines, col = "#ffffff", lwd = 0.75, add = TRUE)
plot(ne_country_lines, col = "#ffffff", lwd = 0.75, add = TRUE)
box()

# Use if saving the plot to PDF
# dev.off()
```

## GAM Prediction Plot

Next, we will plot a map of relative abundance predictions for the white ibis made across BCR 31 using the quasi-Poisson GAM.

```{r}
# Load quasi-Poisson GAM
load("white_ibis_gam.model")

# Predict using the GAM across the study area
pred <- predict(qpois.gam, newdata = pred_surface_eff, 
                type = "link", 
                se.fit = TRUE) %>% 
  as_tibble() %>% 
  # calculate confidence limits and back transform
  transmute(pred_count = qpois.gam$family$linkinv(fit)) %>%
  # add to prediction surface
  bind_cols(pred_surface_eff, .) %>% 
  select(latitude, longitude, pred_count)
```

Convert the dataframe to spatial features, then rasterize the points with the prediction surface raster template:

```{r}
r_pred <- pred %>% 
  # convert to spatial features
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  select(pred_count) %>% 
  st_transform(crs = projection(r)) %>% 
  # rasterize
  rasterize(r)
r_pred <- r_pred[[-1]]

# save the rasters
tif_dir <- "output"
if (!dir.exists(tif_dir)) {
  dir.create(tif_dir)
}
writeRaster(r_pred[["pred_count"]], 
            filename = file.path(tif_dir, "abundance-model_pred-count_white-ibis.tif"),
            overwrite = TRUE)
```

Map white ibis relative abundance across BCR 31.

```{r}
# Use if saving the plots to PDF
# pdf("gam_map.pdf") 

# any expected abundances below this threshold are set to zero
zero_threshold <- 0.05

# Project predictions
r_pred_proj <- projectRaster(r_pred, crs = map_proj$proj4string, method = "ngb")

# par(mfrow = c(2, 1))
for (nm in names(r_pred)) {
  r_plot <- r_pred_proj[[nm]]
  
# Set first margin at 3.5 if knitting file in R Markdown
# Set first margin to 5 if saving plots to PDF
  par(mar = c(3.5, 0.25, 0.25, 0.25))
  # set up plot area
  plot(bcr, col = NA, border = NA)
  plot(ne_land, col = "#dddddd", border = "#888888", lwd = 0.75, add = TRUE)
  
  # modified plasma palette
  plasma_rev <- rev(plasma(25, end = 0.9))
  gray_int <- colorRampPalette(c("#dddddd", plasma_rev[1]))
  pal <- c(gray_int(4)[2], plasma_rev)
  
  # Plot relative abundance predictions
    title <- "White Ibis Relative Abundance"
    # set very low values to zero
    r_plot[r_plot <= zero_threshold] <- NA
    # log transform
    r_plot <- log10(r_plot)
    # breaks and legend
    mx <- ceiling(100 * cellStats(r_plot, max)) / 100
    mn <- floor(100 * cellStats(r_plot, min)) / 100
    brks <- seq(mn, mx, length.out = length(pal) + 1)
    lbl_brks <- seq(mn, mx, length.out = 5)
    lbls <- round(10^lbl_brks, 2)
  
  # abundance plot
  plot(r_plot, 
       col = pal, breaks = brks, 
       maxpixels = ncell(r_plot),
       legend = FALSE, add = TRUE)

  # borders
  plot(bcr, border = "#000000", col = NA, lwd = 1.5, add = TRUE)
  plot(ne_state_lines, col = "#ffffff", lwd = 0.75, add = TRUE)
  plot(ne_country_lines, col = "#ffffff", lwd = 0.75, add = TRUE)
  box()


  # legend
  par(new = TRUE, mar = c(0.5, 0.5, 0.5, 0.5))
  image.plot(zlim = range(brks), legend.only = TRUE, col = pal,
             smallplot = c(0.25, 0.75, 0.06, 0.09),
             horizontal = TRUE,
             axis.args = list(at = lbl_brks, 
                              labels = lbls,
                              fg = "black", col.axis = "black",
                              cex.axis = 0.75, lwd.ticks = 0.5,
                              padj = -1.5),
             legend.args = list(text = title,
                                side = 3, col = "black",
                                cex = 1, line = 0))
}

# Use if saving the plots to PDF
# dev.off()
```

## HGAM Prediction Plot

Finally, we will plot a map of relative abundance predictions for the white ibis made across BCR 31 using the quasi-Poisson HGAM.

We will load the white ibis quasi-Poisson HGAM.

```{r}
# Load HGAM model
load("white_ibis_hgam.model")
```

Create predictions for relative abundance across the prediction surface:

```{r}
# Save a model matrix for the white ibis prediction surface data
# Prediction surface data previously established with the name pred_surface_eff
white.ibis.matrix <- predict(qpois.gam, newdata = pred_surface_eff, type = "lpmatrix")

# Save the vector of coefficient estimates for white.ibis.hgam
white.ibis.coef <- hgam.soi$fixef
white.ibis.coef <- as.matrix(white.ibis.coef)

# Generate "link" predictions
white.ibis.link <- (white.ibis.matrix %*% white.ibis.coef)

# Switch into units of response
white.ibis.pred <- exp(white.ibis.link)
```

Attach predictions to spatial information:

```{r}
pred_2 <- white.ibis.pred %>%
  as.tibble() %>%
  transmute(pred_count = white.ibis.pred) %>%
  bind_cols(pred_surface_eff, .) %>%
  select(latitude, longitude, pred_count)
```

Convert the dataframe to spatial features, then rasterize the points with the prediction surface raster template:

```{r}
r_pred_2 <- pred_2 %>% 
  # convert to spatial features
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>% 
  select(pred_count) %>% 
  st_transform(crs = projection(r)) %>% 
  # rasterize
  rasterize(r)

r_pred_2 <- r_pred_2[[-1]]

# save the rasters
tif_dir_2 <- "output"
if (!dir.exists(tif_dir_2)) {
  dir.create(tif_dir_2)
}
writeRaster(r_pred_2[["pred_count"]], 
            filename = file.path(tif_dir_2, "abundance-model_pred-count_white-ibis_2.tif"),
            overwrite = TRUE)
```

Plot the map of predicted relative abundance across the prediction surface of BCR 31:

```{r}
# Use if saving the plot to PDF
# pdf("hgam_map.pdf") 

# any expected abundances below this threshold are set to zero
zero_threshold <- 0.05

# Project predictions
r_pred_proj_2 <- projectRaster(r_pred_2, crs = map_proj$proj4string, method = "ngb")

  r_plot_2 <- r_pred_proj_2
  
# Set first margin at 3.5 if knitting file in R Markdown
# Set first margin to 5 if saving plots to PDF
  par(mar = c(3.5, 0.25, 0.25, 0.25))
  # set up plot area
  plot(bcr, col = NA, border = NA)
  plot(ne_land, col = "#dddddd", border = "#888888", lwd = 0.75, add = TRUE)
  
  # modified plasma palette
  plasma_rev <- rev(plasma(25, end = 0.9))
  gray_int <- colorRampPalette(c("#dddddd", plasma_rev[1]))
  pal <- c(gray_int(4)[2], plasma_rev)
  
  # abundance, part 1
    title <- "White Ibis Relative Abundance"
    # set very low values to zero
    r_plot_2[r_plot_2 <= zero_threshold] <- NA
    # log transform
    r_plot_2 <- log10(r_plot_2)
    # breaks and legend
    mx <- ceiling(100 * cellStats(r_plot_2, max)) / 100
    mn <- floor(100 * cellStats(r_plot_2, min)) / 100
    brks <- seq(mn, mx, length.out = length(pal) + 1)
    lbl_brks <- seq(mn, mx, length.out = 5)
    lbls <- round(10^lbl_brks, 2)
  
  # abundance, part 2
  plot(r_plot_2, 
       col = pal, breaks = brks, 
       maxpixels = ncell(r_plot_2),
       legend = FALSE, add = TRUE)

  # borders
  plot(bcr, border = "#000000", col = NA, lwd = 1.5, add = TRUE)
  plot(ne_state_lines, col = "#ffffff", lwd = 0.75, add = TRUE)
  plot(ne_country_lines, col = "#ffffff", lwd = 0.75, add = TRUE)
  box()

  # legend
  par(new = TRUE, mar = c(0.5, 0.5, 0.5, 0.5))
  image.plot(zlim = range(brks), legend.only = TRUE, col = pal,
             smallplot = c(0.25, 0.75, 0.06, 0.09),
             horizontal = TRUE,
             axis.args = list(at = lbl_brks, 
                              labels = lbls,
                              fg = "black", col.axis = "black",
                              cex.axis = 0.75, lwd.ticks = 0.5,
                              padj = -1.5),
             legend.args = list(text = title,
                                side = 3, col = "black",
                                cex = 1, line = 0))

# Use if saving the plot to PDF
# dev.off()
```

