---
title: 'Part 6: Investigating Zeros and Nonzeros'
author: "Veronica Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)
library(dggridR)
library(lubridate)
library(ranger)
library(scam)
library(PresenceAbsence)
library(verification)
library(ebirdst)
library(fields)
library(gridExtra)
library(tidyverse)
library(pscl)
library(MASS)
library(numform)

# Spatial packages
library(spdep)
library(rgdal)
library(tripack)
library(dbscan)
library(hglm)

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

## White Ibis

### Load the data

```{r}
# Load training data
ibis <- read_csv("data/ebd_white_ibis_train.csv", show_col_types = FALSE)

# Load test data
ibis.2017 <- read_csv("data/ebd_white_ibis_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(ibis.2017, obs = bird_count)

# Filter test data for zeros
ibis.2017.zeros <- filter(ibis.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(ibis.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
ibis.2017.nonzeros <- filter(ibis.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(ibis.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("white_ibis_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, ibis.2017, type = "response") %>%
  tibble(family = "White Ibis GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, ibis.2017.zeros, type = "response") %>%
  tibble(family = "White Ibis GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, ibis.2017.nonzeros, type = "response") %>%
  tibble(family = "White Ibis GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("white_ibis_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = ibis.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "White Ibis HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = ibis.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "White Ibis HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = ibis.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "White Ibis HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Glossy Ibis

## Clear the workspace

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
ibis <- read_csv("data/ebd_glossy_ibis_train.csv", show_col_types = FALSE)

# Load test data
ibis.2017 <- read_csv("data/ebd_glossy_ibis_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(ibis.2017, obs = bird_count)

# Filter test data for zeros
ibis.2017.zeros <- filter(ibis.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(ibis.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
ibis.2017.nonzeros <- filter(ibis.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(ibis.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("glossy_ibis_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, ibis.2017, type = "response") %>%
  tibble(family = "Glossy Ibis GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, ibis.2017.zeros, type = "response") %>%
  tibble(family = "Glossy Ibis GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, ibis.2017.nonzeros, type = "response") %>%
  tibble(family = "Glossy Ibis GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("glossy_ibis_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = ibis.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Glossy Ibis HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = ibis.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Glossy Ibis HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = ibis.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Glossy Ibis HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Great Egret

## Clear the workspace

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
egret <- read_csv("data/ebd_great_egret_train.csv", show_col_types = FALSE)

# Load test data
egret.2017 <- read_csv("data/ebd_great_egret_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(egret.2017, obs = bird_count)

# Filter test data for zeros
egret.2017.zeros <- filter(egret.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(egret.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
egret.2017.nonzeros <- filter(egret.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(egret.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("great_egret_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, egret.2017, type = "response") %>%
  tibble(family = "Great Egret GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, egret.2017.zeros, type = "response") %>%
  tibble(family = "Great Egret GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, egret.2017.nonzeros, type = "response") %>%
  tibble(family = "Great Egret GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("great_egret_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = egret.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Great Egret HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = egret.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Great Egret HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = egret.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Great Egret HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Cattle Egret

## Clear the workspace

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
egret <- read_csv("data/ebd_cattle_egret_train.csv", show_col_types = FALSE)

# Load test data
egret.2017 <- read_csv("data/ebd_cattle_egret_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(egret.2017, obs = bird_count)

# Filter test data for zeros
egret.2017.zeros <- filter(egret.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(egret.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
egret.2017.nonzeros <- filter(egret.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(egret.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("cattle_egret_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, egret.2017, type = "response") %>%
  tibble(family = "Cattle Egret GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, egret.2017.zeros, type = "response") %>%
  tibble(family = "Cattle Egret GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, egret.2017.nonzeros, type = "response") %>%
  tibble(family = "Cattle Egret GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("cattle_egret_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = egret.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Cattle Egret HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = egret.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Cattle Egret HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = egret.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Cattle Egret HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Snowy Egret

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
egret <- read_csv("data/ebd_snowy_egret_train.csv", show_col_types = FALSE)

# Load test data
egret.2017 <- read_csv("data/ebd_snowy_egret_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(egret.2017, obs = bird_count)

# Filter test data for zeros
egret.2017.zeros <- filter(egret.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(egret.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
egret.2017.nonzeros <- filter(egret.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(egret.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("snowy_egret_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, egret.2017, type = "response") %>%
  tibble(family = "Snowy Egret GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, egret.2017.zeros, type = "response") %>%
  tibble(family = "Snowy Egret GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, egret.2017.nonzeros, type = "response") %>%
  tibble(family = "Snowy Egret GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("snowy_egret_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = egret.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Snowy Egret HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = egret.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Snowy Egret HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = egret.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Snowy Egret HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Great Blue Heron

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
heron <- read_csv("data/ebd_gbh_train.csv", show_col_types = FALSE)

# Load test data
heron.2017 <- read_csv("data/ebd_gbh_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(heron.2017, obs = bird_count)

# Filter test data for zeros
heron.2017.zeros <- filter(heron.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(heron.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
heron.2017.nonzeros <- filter(heron.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(heron.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("gbh_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, heron.2017, type = "response") %>%
  tibble(family = "Great Blue Heron GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, heron.2017.zeros, type = "response") %>%
  tibble(family = "Great Blue Heron GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, heron.2017.nonzeros, type = "response") %>%
  tibble(family = "Great Blue Heron GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("gbh_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = heron.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Great Blue Heron HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = heron.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Great Blue Heron HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = heron.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Great Blue Heron HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Little Blue Heron

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
heron <- read_csv("data/ebd_lbh_train.csv", show_col_types = FALSE)

# Load test data
heron.2017 <- read_csv("data/ebd_lbh_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(heron.2017, obs = bird_count)

# Filter test data for zeros
heron.2017.zeros <- filter(heron.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(heron.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
heron.2017.nonzeros <- filter(heron.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(heron.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("lbh_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, heron.2017, type = "response") %>%
  tibble(family = "Little Blue Heron GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, heron.2017.zeros, type = "response") %>%
  tibble(family = "Little Blue Heron GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, heron.2017.nonzeros, type = "response") %>%
  tibble(family = "Little Blue Heron GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("lbh_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = heron.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Little Blue Heron HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = heron.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Little Blue Heron HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = heron.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Little Blue Heron HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

## Green Heron

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
as_factor <- forcats::as_factor
```

### Load the data

```{r}
# Load training data
heron <- read_csv("data/ebd_green_heron_train.csv", show_col_types = FALSE)

# Load test data
heron.2017 <- read_csv("data/ebd_green_heron_2017.csv", show_col_types = FALSE)
# and select for only the count values
obs.count <- select(heron.2017, obs = bird_count)

# Filter test data for zeros
heron.2017.zeros <- filter(heron.2017, bird_count == 0)
# and select for only the count values
obs.count.zeros <- select(heron.2017.zeros, obs = bird_count)

# Filter test data for nonzeros
heron.2017.nonzeros <- filter(heron.2017, bird_count > 0)
# and select for only the count values
obs.count.nonzeros <- select(heron.2017.nonzeros, obs = bird_count)
```

### Generate GAM predictions

```{r}
# Load the quasi-Poisson GAM
load("green_heron_gam.model")

# Generate GAM predictions over full test set
gam.pred1 <- predict(qpois.gam, heron.2017, type = "response") %>%
  tibble(family = "Green Heron GAM, all", pred = .) %>%
  bind_cols(obs.count)

# Generate predictions over zeros
gam.pred2 <- predict(qpois.gam, heron.2017.zeros, type = "response") %>%
  tibble(family = "Green Heron GAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)

# Generate predictions over nonzeros
gam.pred3 <- predict(qpois.gam, heron.2017.nonzeros, type = "response") %>%
  tibble(family = "Green Heron GAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Generate HGAM predictions, full test set

```{r}
# Load the quasi-Poisson HGAM
load("green_heron_hgam.model")

# Save a model matrix for 2017 test data
matrix.17 <- predict(qpois.gam, newdata = heron.2017, type = "lpmatrix")

# Save the vector of coefficient estimates
hgam.coef <- as.matrix(hgam.soi$fixef)

# Generate predictions (link)
link.17 <- (matrix.17 %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred1 <- as.vector(pred.17) %>%
  tibble(family = "Green Heron HGAM, all", pred = .) %>%
  bind_cols(obs.count)
```

### Generate HGAM predictions, zeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.zeros <- predict(qpois.gam, newdata = heron.2017.zeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.zeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred2 <- as.vector(pred.17) %>%
  tibble(family = "Green Heron HGAM, zeros", pred = .) %>%
  bind_cols(obs.count.zeros)
```

### Generate HGAM predictions, nonzeros only

```{r}
# Save a model matrix for 2017 test data
matrix.17.nonzeros <- predict(qpois.gam, newdata = heron.2017.nonzeros, type = "lpmatrix")

# Generate predictions (link)
link.17 <- (matrix.17.nonzeros %*% hgam.coef)

# Put in units of the response
pred.17 <- exp(link.17)

# Build tibble with predictions and observed counts
hgam.pred3 <- as.vector(pred.17) %>%
  tibble(family = "Green Heron HGAM, nonzeros", pred = .) %>%
  bind_cols(obs.count.nonzeros)
```

### Calculate MAD

```{r}
# Bind the predictions
predictions <- bind_rows(gam.pred1, hgam.pred1, gam.pred2, hgam.pred2, gam.pred3, hgam.pred3) %>%
  mutate(family = as_factor(family))
```

```{r}
# Calculate MAD and corresponding SD for each set of predictions
mad.tibble <- predictions %>% 
  group_by(family) %>% 
  summarise(mad = mean(abs(obs - pred), na.rm = TRUE),
            sd = sd(abs(obs - pred))) %>% 
  ungroup()

# Print the table
mad.tibble
```

