---
title: "Part 4.1: Moran's I Test for Spatial Autocorrelation"
author: "Veronica Lee"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(sf)
library(raster)
library(lubridate)
library(ranger)
library(scam)
library(PresenceAbsence)
library(verification)
library(ebirdst)
library(fields)
library(gridExtra)
library(tidyverse)
library(pscl)
library(MASS)
library(numform)

# Spatial packages
library(spdep)
library(rgdal)
library(tripack)
library(dbscan)
library(hglm)

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

## White Ibis

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_white_ibis_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the white ibis
load("white_ibis_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
nrow(bird)

nrow(unique(data.frame(bird$latitude, bird$longitude)))
```

In our white ibis dataset, there are 3787 observations, but only 1860 unique coordinate points. 

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

This p-value is extremely small, close to zero. We do have spatial autocorrelation in the case of the white ibis.

## Glossy Ibis

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_glossy_ibis_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the glossy ibis
load("glossy_ibis_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

## Great Egret

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_great_egret_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the great egret
load("great_egret_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

## Cattle Egret

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_cattle_egret_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the cattle egret
load("cattle_egret_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

## Snowy Egret

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_snowy_egret_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the snowy egret
load("snowy_egret_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

## Great Blue Heron

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_gbh_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the great blue heron
load("gbh_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

## Little Blue Heron

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_lbh_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the little blue heron
load("lbh_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

## Green Heron

```{r}
# Clear the workspace
rm(list=ls())

# Resolve namespace conflicts
select <- dplyr::select
map <- purrr::map
projection <- raster::projection
```

Prepare the data

```{r}
# Load the training data
bird <- read_csv("data/ebd_green_heron_train.csv", show_col_types = FALSE)
```

```{r}
# Load the quasi-Poisson GAM for the green heron
load("green_heron_gam.model")

# Extract the residuals (in units of the response)
resid <- residuals(qpois.gam, type = "response")

# Check that the length of our residual vector matches with the row dimension of the bird dataset
length(resid) == nrow(bird)

# Attach the residuals to the data
bird <- bird %>% mutate(resid = resid)
```

```{r}
# Create a "lat-long" ID code
bird <- bird %>% mutate(paste(bird$latitude, bird$longitude, sep = "_"))

# Rename the new variable
bird <- rename(bird, latlong_id = `paste(bird$latitude, bird$longitude, sep = "_")`)

# If it's worked correctly, this will output TRUE
length(unique(bird$latlong_id)) == 1860
```

Now what we want to do is find the average (median) residual value for each latlong ID. (This is because Moran's I won't run for datasets with multiple observations in the same coordinate location.)

```{r}
avg_resid_df <- bird %>%
  group_by(latlong_id) %>%
  summarize(avg_resid = median(resid),
            latitude = latitude,
            longitude = longitude)

# Generate a subset of the above dataset that has one observation per location
unique.latlong <- do.call(rbind, by(avg_resid_df, avg_resid_df$latlong_id,
                  FUN = function(x) head(x, 1)))

nrow(unique.latlong)

head(unique.latlong)
```

Now our goal is to run Moran's I test on the residuals from the first subsample.

```{r}
coords.m <- as.matrix(cbind(unique.latlong$latitude, unique.latlong$longitude))
IDs <- unique.latlong$latlong_id

# First, create a Delaunay neighbor object
del.nb <- tri2nb(coords.m, row.names = IDs)

# Then, from that, a Sphere of Influence neighbor object
soi.nb <- graph2nb(soi.graph(del.nb, coords.m), row.names = IDs)
```

```{r}
# Check the assumption of continuity for our residuals
ggplot(data = unique.latlong, mapping = aes(x = avg_resid)) +
  geom_histogram(bins = 100)
```

Good enough!

Now we will run our Moran's I test

```{r}
moran <- moran.test(unique.latlong$avg_resid, listw = nb2listw(soi.nb, style = "B"))

moran$p.value

round2(moran$p.value, 3)
```

-End